### 2020-10-14 reports/article.tex
monitoring stations figure
addressing Remi's comments

Urbanization is nowadays a global phenomenon that is increasingly concentrating the world's population in urban areas, which are expected to grow in both the number of dwellers and spatial extent \cite{seto2011meta, angel2012atlas, unitednations2018world}.
As a major force of landscape change, urbanization is characterized by the conversion of natural to artificial surfaces, which alters the energy and water exchanges as well as the movement of air. Such changes often result in the urban heat island (UHI) effect, a phenomena by which urban temperatures are warmer than its rural surroundings \cite{oke1982energetic}.

% As urban areas grow and global temperatures increase, heat waves are expected to become more intense and frequent ant the UHI effect
As urban areas grow and global temperatures rise, the UHI effect is expected to become more intense \cite{meehl2004more,huang2019projecting}, which makes urban heat mitigation a major priority for urban planning and policy-making.
% Green infrastructure
% To that end, cities are increasingly
Increasing urban green space, especially the urban tree canopy, has been one of the most widely advocated strategies of urban heat mitigation \cite{bowler2010urban,phelan2015urban,koc2018evaluating}.
% Overall decrease of urban green space in European cities due to infill development \cite{haaland2015challenges}.
% TODO: challanges with densification, trade-offs in the provision of ecosystem services
% TODO: need for methods to quantify
Nevertheless, the impact of the abundance and spatial configuration of the urban tree canopy on air temperature shows a complex spatial behaviour that remains
Therefore, there is an urgent need to develop robust and spatially explicit methods to quantify the contribution of urban green space to urban heat mitigation \cite{haaland2015challenges,artmann2019urban}.

The present work introduces a novel approach to spatially evaluate the heat mitigation potential that could be achieved by increasing the urban tree canopy cover in the urban agglomeration of Lausanne, Switzerland. The proposed method consists of three major parts.
% 1. Couple the cadastral LULC map with a high-resolution tree canopy map to detect locations where the tree canopy cover can be increased
% 2. Generate a variety of synthetic scenarios by increasing the tree canopy cover, which can mimic urban greening strategies. % in different spatial configurations
% 3. 
First, the cadastral LULC map is coupled with a high-resolution tree canopy map to detect candidate pixels where the tree canopy cover can be increased.
Secondly, a set of synthetic scenarios is generated by increasing the tree canopy cover in the candidate pixels.
% is applied with the LULC map that corresponds to 
% Finally, the synthetic LULC map of each scenario is fed to the InVEST urban cooling model in order to estimate the air temperature of
Finally, the spatial distribution of air temperature of each scenario is estimated by applying the InVEST urban cooling model to the corresponding synthetic LULC map.

@@

Lausanne is the fourth largest Swiss urban agglomeration with 420757 inhabitants as of January 2019 \cite{sfso2018city}. The agglomeration is situated at the Swiss Plateau and on the shore of the Lake L\'eman, and is characterized by a continental temperate climate with mean annual temperatures of 10.9 $\degree$C and mean annual precipitation of 100 mm, with a dominating vegetation of mixed broadleaf forest. The spatial extent of the study has been selected following the recent application of the InVEST urban cooling model to Lausanne by Bosch et al. \cite{bosch2020spatially}, and covers an area of 112.46 $km^2$.

In order to evaluate the human thermal comfort, the population data for the study area has been extracted from the population and households statistics (STATPOP) \cite{sfso2020statistique} provided by the Swiss Federal Statistical Office (SFSO) with the Python library swisslandstats-geopy \cite{bosch2019swisslandstats}.

@@

\begin{itemize}
\item \texttt{lucode} the LULC class code as represented in the LULC raster map
\item \texttt{Shade} a value between 0 and 1 representing the proportion of tree cover in such LULC class
\item \texttt{Kc} the evapotranspiration coefficient
\item \texttt{Albedo} a value between 0 and 1 representing the proportion of solar radiation directly reflected by the LULC class
\item \texttt{Green\_area} whether the LULC class should be considered a green area
% \item \texttt{Building\_intensity} a value between 0 and 1 representing the ratio between floor area and land area
\end{itemize}

In order to simulate the spatial distribution of $T_{air}$, the model requires two additional inputs. The first is the rural reference temperature $T_{ref}$, where the UHI effect is not observed, e.g., in the rural surroundings of the city. The second is the magnitude of the urban heat island effect $UHI_{max}$, namely the difference between the rural refrence temperature and the maximum $T_{air}$ observed in the city center.

@@

% To better account for the spatial heterogeneity of cities, a procedure to redefine the LULC classes from the cadastral survey has been designed, which further distinguishes the LULC classes depending on their proportional cover of both trees and buildings. Such reclassification is based on combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
A procedure to redefine the LULC classes from the cadastral survey has been designed to distinguish the LULC classes depending on their proportional cover of both trees and buildings. The reclassification is achieved by combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
The 1~m binary tree canopy mask has been derived from the SWISSIMAGE orthomosaic \cite{swisstopo2019swissimage}, by means of the Python library DetecTree \cite{bosch2020detectree}, which implements the methods proposed by Yang et al. \cite{yang2009tree}. On the other hand, the 1~m binary building mask has been obtained by rasterizing the buildings of the vector cadastral survey \cite{asitvd2018structure}.

@@

\subsubsection*{Simulation of urban greening by randomly increasing the proportion of tree cover}
With the aim of evaluating the effect of the proportion of tree cover on air temperature, a set of scenarios is generated by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy cover can be increased.
For each of the aforementioned percentages, 10 scenarios will be generated by randomly sampling from the candidate pixels until the desired proportion of changed pixels is matched. Therefore, the temperature maps for each of the percentage of changed pixels will be computed by averaging the 10 respective simulated runs.

\subsubsection*{Simulation of distinct urban greening approaches by clustering and scattering trees}

% To expore the relationship between
In order to obtain a greater variety of spatial patterns of tree canopy cover and further explore the relationship between the spatial configuration of tree canopy and UHIs, two additional sets of scenarios will be generated.
% Like the first set of scenarios described above, these two additional sets are also produced by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy can be increased.
% Nevertheless, in these two sets the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., > 75\%) found in their Moore neighborhood. Therefore,
In these sets of scenarios, the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., greater than 75\%) found in their Moore neighborhood.
Therefore, the first set of scenarios pixels with higher number of high tree canopy cover neighbors are transformed first, which intends to spatially cluster pixels of high tree canopy cover. Conversely, the second set of scenarios intends to spatially scatter pixels of high tree canopy cover by giving precedence to pixels with lower number of high tree canopy neighbors.
To explore the difference between the two approaches with more detail, these two sets of scenarios will be produced by transforming a 12.5, 25, 37.5, 50, 62.5, 75, 87.5 and 100\% of the pixels where the tree canopy can be increased.
Like the previous set of scenarios, the temperature maps will be computed by averaging the results of 10 simulation funs for each configuration, i.e., combination of proportion of changed pixels and spatial approach of either clustering or scattering the transformed pixels.

@@

The scenarios simulated in this study represent a new form of exploring the heat mitigation potential provided by the urban tree canopy cover.
The results map locations where the tree canopy cover in the urban agglomeration of Lausanne can be increased, and shows that such changes can result in urban nighttime temperatures that are up to 2$\degree$C lower.
Additionally, the number of dwellers exposed to nighttime temperatures over 24$\degree$C can be reduced from 217051 to 31905, which respectively represents a 78.40 and 11.52\% of the total population in the study area.

% 1. EFFECTS OF THE SPATIAL CONFIGURATION
% % The proposed simulation approach presents several major advantages when compared to statistical analyses.
% Overall, the simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, both in terms of the average temperature and the number of urban dwellers exposed to the highest tempetratures.
% % might have the spatial configuration of the tree canopy might indeed have a significant effect on the spatial pattern of air temperature, yet it seems that such an effect is very small when compared to the effect of the proportion of tree cover
% Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% % Trees contribute to urban heat mitigation in two primary ways, namely the shading and evapotranspiration processes. Therefore, the effect of the spatial configuraiton of trees on its cooling effectiveness
% The effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes. Such a relationship is know to be highly complex and dependent on the ecological context \cite{zhou2017effects}, tree species \cite{jiao2017patch}
The simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, which is in line with previous studies in humid climates \cite{kong2014effects,estoque2017effects,zhou2017effects,yu2018strong,nastran2019urban}.
Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes, which is known to be highly complex relationship that is strongly mediated by the tree species as well as the local climatic and environmental conditions \cite{zhou2017effects,jiao2017patch,yu2018strong,wang2020significant}. Furthermore, the cooling effects of tree canopy cover are scale-dependent \cite{li2013relationship,estoque2017effects,yan2019testing,wang2020significant,terfa2020spatial}.
Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes.
Such a relationship is known to be strongly mediated by the tree species, background climatic and environmental conditions as well as the spatial scale \cite{li2013relationship,estoque2017effects,zhou2017effects,jiao2017patch,yu2018strong,yan2019testing,wang2020significant,terfa2020spatial}.
In order to better represent the effects of the spatial configuration of the tree canopy, the InVEST urban cooling model must be further validated with experiments at the neighborhood scale to ensure that it provides an appropriate city-scale depiction of how the urban heat mitigation mechanisms operate at finer scales.

@@

% 3. EVALUATE FUTURE SCENARIOS, COMPARE SPRAWL VS INFILL
% Additionally, we can use the approach to compare distinct urban expansion scenarios (e.g., compact, sprawl) \cite{lemonsu2015vulnerability,yang2016contrasting}, also landscape homogenization along the urban-rural gradient \cite{wang2020spatial}.
Despite the limitations noted above, a major advantage of the proposed approach is that it can be used to evaluate urban heat mitigation of synthetic scenarios.
The simulations presented in this article focus on spatially exploring the effects of an increase of the tree canopy cover, yet there is room for much more experimentation of this kind.
% MAYBE: landscape ecology principals/network planning to ensure connectivity and the like? see \cite[page 766]{haaland2015challenges}
For instance, in line with recent studies \cite{lemonsu2015vulnerability,yang2016contrasting,trimmel2019thermal}, the approach could be applied to examine the impact of distinct urbanization scenarios such as densification and urban sprawl on air temperature and human thermal comfort, under current conditions as well as future climate estimates, e.g., by changing the $T_{ref}$ or $UHI_{max}$ parameters.
Similarly, InVEST urban cooling model might be coupled with models of LULC change such as cellular automata in order to assess not only which scenarios are most desirable in terms of urban heat mitigation, but also which planning strategies might lead to them \cite{silva2008strategies,bosch2019addressing}.
% Also to evaluate projected LULC maps (e.g., outputs of cellular-automata simulations \cite{bosch2019addressing})
The spatiotemporal patterns of LULC change observed during the last 40 years in the Lausanne agglomeration have been characterized by infilling development and a progressive coalescensce of the artificial surfaces in its inner ring \cite{bosch2020spatiotemporal}.
% Overall decrease of urban green space in European cities due to infill development \cite{haaland2015challenges}.
Therefore, such an infilling trend urges for careful evaluation of the beneficial ecosystem services provided by urban green spaces, which should be balanced against the adverse consequences of urban sprawl \cite{haaland2015challenges,artmann2019urban}.

% \section*{Conclusion}
% TODO: conclusions

@@

\begin{figure}[ht]
  \centering
  % \includegraphics[width=.98\textwidth]{figures/monitoring-stations}
  \caption{\label{fig:monitoring-stations} Locations of the monitoring stations used to get the $T_{ref}$ and $UHI_{max}$ parameters. The basemap tile is provided by StamenDesign, under CC BY 3.0, with data from OpenStreetMap, under ODbL.}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-regplot}
  \caption{\label{fig:scenarios-prop-regplot} Relationship between the proportion of candidate pixels transformed and the average simulated temperature $\overline{T}$ for each scenario sample. The translucent bands around the regression line represent the 95\% confidence intervals estimated using a bootstrap.}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-hists}
  \caption{\label{fig:scenarios-prop-hists} Histogram of raster temperature values for a 25, 50, 75 and 100\% of the candidate pixels transformed. The temperature rasters for each histogram are computed by averaging the 10 simulations with the same proportion of candidate pixels transformed.}
\end{figure}

>>>>

Urbanization is nowadays a global phenomenon that is increasingly concentrating the world's population in urban areas, which are expected to grow in both the number of dwellers and spatial extent \cite{seto2011meta, angel2012atlas, unitednations2018world}.
As a major force of landscape change, urbanization is characterized by the conversion of natural to artificial surfaces, which alters the energy and water exchanges as well as the movement of air. Such changes often result in the urban heat island (UHI) effect, a phenomenon by which urban temperatures are warmer than its rural surroundings \cite{oke1973city,oke1982energetic,arnfield2003two,voogt2003thermal,grimmond2007urbanization,phelan2015urban}. % \cite{oke1982energetic}.

% The negative impacts of UHI have been widely documented and include increased water and energy consumption \cite{santamouris2015impact}, reduced productivity \cite{kjellstrom2009workplace} and increased health risks \cite{kovats2008heat,laaidi2012impact}.
The negative impacts of UHI have been widely documented and include increased energy and water consumption \cite{akbari2001cool,golden2006energy,santamouris2015impact}, reduced workplace productivity \cite{kjellstrom2009workplace,zander2015heat} and aggravation of health risks \cite{chestnut1998analysis,kovats2008heat,laaidi2012impact}.
% As urban areas grow and global temperatures increase, heat waves are expected to become more intense and frequent ant the UHI effect
As urban areas grow and global temperatures rise, the UHI effect is expected to become more intense \cite{meehl2004more,huang2019projecting}, which makes urban heat mitigation a major priority for urban planning and policy-making.
% Green infrastructure
% To that end, cities are increasingly
Increasing urban green space, especially the urban tree canopy, has been one of the most widely advocated strategies of urban heat mitigation, nevertheless, the impacts of the urban tree canopy on air temperature show a complex spatial behaviour that remains poorly understood \cite{bowler2010urban,phelan2015urban,koc2018evaluating}.
Therefore, there is an urgent need to develop robust and spatially explicit methods to quantify the contribution of urban green space to urban heat mitigation \cite{haaland2015challenges,artmann2019urban}.

The present work introduces a novel approach to spatially evaluate the heat mitigation potential that could be achieved by increasing the urban tree canopy cover. The proposed method consists of three major parts.
% 1. Couple the cadastral LULC map with a high-resolution tree canopy map to detect locations where the tree canopy cover can be increased
% 2. Generate a variety of synthetic scenarios by increasing the tree canopy cover, which can mimic urban greening strategies. % in different spatial configurations
% 3. 
First, the cadastral LULC map is coupled with a high-resolution tree canopy map to detect candidate pixels where the tree canopy cover can be increased.
Secondly, a synthetic scenario is generated by increasing the tree canopy cover in a fraction of the candidate pixels --- the quantity of transformed pixels can be set exogenously, e.g., to mimic available budgets and resources.
The transformed pixels can be selected from the candidate pixels by randomly sampling, or pursuing specific spatial configurations by scattering or clustering them to other pixels with high tree canopy cover.
% is applied with the LULC map that corresponds to 
% Finally, the synthetic LULC map of each scenario is fed to the InVEST urban cooling model in order to estimate the air temperature of
Finally, the spatial distribution of air temperature of each scenario is estimated by applying the InVEST urban cooling model to the corresponding synthetic LULC map.
By applying such a procedure in the urban agglomeration of Lausanne, Switzerland, this study aims to map the heat mitigation potential that can be achieved departing from the existing urban fabric.
In order to quantify the effects of the abundance and spatial configuration of the tree canopy cover on urban heat mitigation, a set of synthetic scenarios are generated by increasing different proportions of tree canopy cover in distinct spatial configurations. % and human thermal comfort
% The latter is achieved by coupling a gridded population census with the temperature maps simulated with different proportions and spatial configurations of tree canopy cover.
Additionally, the human thermal comfort of each scenario is evaluated by coupling the simulated temperature map with a gridded population census.
% TODO: the aim is to do something that helps planning
% The results

@@

Lausanne is the fourth largest Swiss urban agglomeration with 420757 inhabitants as of January 2019 \cite{sfso2018city}. The agglomeration is located at the Swiss Plateau and on the shore of the Lake L\'eman, and is characterized by a continental temperate climate with mean annual temperatures of 10.9 $\degree$C and mean annual precipitation of 100 mm, with a dominating vegetation of mixed broadleaf forest. The spatial extent of the study has been selected following the recent application of the InVEST urban cooling model to Lausanne by Bosch et al. \cite{bosch2020spatially}, and covers an area of 112.46 $km^2$.

In order to evaluate the human thermal comfort, the population data for the study area has been extracted from the population and households statistics (STATPOP) \cite{sfso2020statistique} provided at a 100~m resolution by the Swiss Federal Statistical Office (SFSO) with the Python library swisslandstats-geopy \cite{bosch2019swisslandstats}.

@@

\begin{itemize}
\item \texttt{lucode} the LULC class code as represented in the LULC raster map
\item \texttt{Shade} a value between 0 and 1 representing the proportion of tree canopy cover in such LULC class
\item \texttt{Kc} the evapotranspiration coefficient
\item \texttt{Albedo} a value between 0 and 1 representing the proportion of solar radiation directly reflected by the LULC class
\item \texttt{Green\_area} whether the LULC class should be considered a green area
% \item \texttt{Building\_intensity} a value between 0 and 1 representing the ratio between floor area and land area
\end{itemize}

In order to simulate the spatial distribution of air temperature ($T_{air}$), the model requires two additional inputs. The first is the rural reference temperature $T_{ref}$, where the UHI effect is not observed, e.g., in the rural surroundings of the city. The second is the magnitude of the urban heat island effect $UHI_{max}$, namely the difference between the rural refrence temperature and the maximum $T_{air}$ observed in the city center.

@@

% To better account for the spatial heterogeneity of cities, a procedure to redefine the LULC classes from the cadastral survey has been designed, which further distinguishes the LULC classes depending on their proportional cover of both trees and buildings. Such reclassification is based on combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
A procedure to redefine the LULC classes from the cadastral survey has been designed to distinguish the LULC classes depending on their proportional cover of both trees and buildings. The reclassification is achieved by combining the 10~m raster LULC map with two 1~m binary raster masks, one for the tree canopy raster and another for the buildings. % , and consists of three main steps.
The 1~m binary tree canopy mask has been derived from the SWISSIMAGE orthomosaic \cite{swisstopo2019swissimage}, by means of the Python library DetecTree \cite{bosch2020detectree}, which implements the methods proposed by Yang et al. \cite{yang2009tree}. The estimated classification accuracy of the tree canopy classification is of 91.75\%. On the other hand, the 1~m binary building mask has been obtained by rasterizing the buildings of the vector cadastral survey \cite{asitvd2018structure}.

@@

\subsubsection*{Simulation of urban greening by randomly increasing the proportion of tree cover}
With the aim of evaluating the effect of the proportion of tree cover on air temperature, a set of scenarios is generated by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy cover can be increased.
For each of the aforementioned percentages, a number scenario runs will be generated by randomly sampling from the candidate pixels until the desired proportion of changed pixels is matched. Therefore, the temperature maps for each of the percentage of changed pixels will be computed by averaging the respective simulated runs.
After observing little variability among the simulation runs, the number of scenario runs has been set to 10.

\subsubsection*{Simulation of distinct urban greening approaches by clustering and scattering the tree canopy}

% To expore the relationship between
In order to obtain a greater variety of spatial patterns of tree canopy cover and further explore the relationship between the spatial configuration of tree canopy and UHIs, two additional sets of scenarios will be generated.
% Like the first set of scenarios described above, these two additional sets are also produced by transforming a 25, 50, 75 and 100\% of the pixels where the tree canopy can be increased.
% Nevertheless, in these two sets the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., > 75\%) found in their Moore neighborhood. Therefore,
In these sets of scenarios, the transformed pixels are not randomly sampled but selected according to the number of pixels with high tree canopy cover (i.e., greater than 75\%) found in their Moore neighborhood.
Therefore, the first set of scenarios pixels with higher number of high tree canopy cover neighbors are transformed first, which intends to spatially cluster pixels of high tree canopy cover. Conversely, the second set of scenarios intends to spatially scatter pixels of high tree canopy cover by giving precedence to pixels with lower number of high tree canopy neighbors.
To explore the difference between the two approaches with more detail, these two sets of scenarios will be produced by transforming a 12.5, 25, 37.5, 50, 62.5, 75 and 87.5 of the pixels where the tree canopy can be increased.
The scenario where all the candidate pixels, i.e., 100\%, are transformed is excluded here because under such such circumstances the clustering or scattering criteria to select the pixels does not have any effect.
Like the previous set of scenarios, the temperature maps will be computed by averaging the results of 10 simulation funs for each configuration, i.e., combination of proportion of changed pixels and spatial approach of either clustering or scattering the transformed pixels. This results in a total of 140 scenarios, where 70 correspond to the clustering approach and the other 70 to the scattering approach.

@@

% \subsection*{Complexity of the impact of the spatial pattern of tree canopy on urban heat mitigation}
\subsection*{Validity and applicability of the proposed approach}

% The scenarios simulated in this study represent a new form of exploring the heat mitigation potential provided by the urban tree canopy cover.
% The results map locations where the tree canopy cover in the urban agglomeration of Lausanne can be increased, and shows that such changes can result in urban nighttime temperatures that are up to 2$\degree$C lower.
% Additionally, the number of dwellers exposed to nighttime temperatures over 24$\degree$C can be reduced from 217051 to 31905, which respectively represents a 78.40 and 11.52\% of the total population in the study area.
The scenarios simulated in this study map locations where the tree canopy cover in the urban agglomeration of Lausanne can be increased, and suggests that such changes can result in urban nighttime temperatures that are up to 2$\degree$C lower.
% 1. EFFECTS OF THE SPATIAL CONFIGURATION
% % The proposed simulation approach presents several major advantages when compared to statistical analyses.
% Overall, the simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, both in terms of the average temperature and the number of urban dwellers exposed to the highest tempetratures.
% % might have the spatial configuration of the tree canopy might indeed have a significant effect on the spatial pattern of air temperature, yet it seems that such an effect is very small when compared to the effect of the proportion of tree cover
% Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% % Trees contribute to urban heat mitigation in two primary ways, namely the shading and evapotranspiration processes. Therefore, the effect of the spatial configuraiton of trees on its cooling effectiveness
% The effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes. Such a relationship is know to be highly complex and dependent on the ecological context \cite{zhou2017effects}, tree species \cite{jiao2017patch}
% The simulated scenarios in Lausanne indicate that given the same proportion of tree canopy cover, a scattered configuration might result in more effective urban heat mitigation than a clustered one, which is in line with previous studies in humid climates \cite{kong2014effects,estoque2017effects,zhou2017effects,yu2018strong,nastran2019urban}.
The results indicate that given the same proportion of tree canopy cover, a scattered configuration might lead to more effective urban heat mitigation than a clustered one, which is in line with previous studies in humid climates \cite{zhou2011does,kong2014effects,estoque2017effects,zhou2017effects,yu2018strong,nastran2019urban}.
Nevertheless, the results suggest that effect of the spatial configuration is secondary when compared to the effect of the composition, i.e., proportion of tree cover.
% Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes, which is known to be highly complex relationship that is strongly mediated by the tree species as well as the local climatic and environmental conditions \cite{zhou2017effects,jiao2017patch,yu2018strong,wang2020significant}. Furthermore, the cooling effects of tree canopy cover are scale-dependent \cite{li2013relationship,estoque2017effects,yan2019testing,wang2020significant,terfa2020spatial}.
Overall, the effect of the spatial configuration of trees on its urban cooling effectiveness depends on how it affects the shading and evapotranspiration processes.
Such a relationship is known to be strongly mediated by the tree species, background climatic and environmental conditions as well as the spatial scale \cite{li2013relationship,estoque2017effects,zhou2017effects,jiao2017patch,yu2018strong,yan2019testing,wang2020significant,terfa2020spatial}.
In order to better represent the effects of the spatial configuration of the tree canopy, the InVEST urban cooling model must be further validated with experiments at the neighborhood scale to ensure that it provides an appropriate city-scale depiction of how the urban heat mitigation mechanisms operate at finer scales.

@@

% \subsection*{Applicability of the proposed approach}

% 3. EVALUATE FUTURE SCENARIOS, COMPARE SPRAWL VS INFILL
% Additionally, we can use the approach to compare distinct urban expansion scenarios (e.g., compact, sprawl) \cite{lemonsu2015vulnerability,yang2016contrasting}, also landscape homogenization along the urban-rural gradient \cite{wang2020spatial}.
Despite the limitations noted above, a major advantage of the proposed approach is that it can be used to evaluate urban heat mitigation of synthetic scenarios.
The simulations presented in this article focus on spatially exploring the effects of an increase of the tree canopy cover, yet there is room for much more experimentation of this kind.
% MAYBE: landscape ecology principals/network planning to ensure connectivity and the like? see \cite[page 766]{haaland2015challenges}
For instance, in line with recent studies \cite{lemonsu2015vulnerability,yang2016contrasting,trimmel2019thermal}, the approach could be applied to examine the impact of distinct urbanization scenarios such as densification and urban sprawl on air temperature and human thermal comfort, under current conditions as well as future climate estimates, e.g., by changing the $T_{ref}$ or $UHI_{max}$ parameters.
Similarly, InVEST urban cooling model might be coupled with models of LULC change such as cellular automata in order to assess not only which scenarios are most desirable in terms of urban heat mitigation, but also which planning strategies might lead to them \cite{silva2008strategies,white2015modeling,bosch2019addressing}.
% Also to evaluate projected LULC maps (e.g., outputs of cellular-automata simulations \cite{bosch2019addressing})

% TODO: multi-scale \cite{artmann2019urban,wang2020spatial}, complimentary to other approaches (e.g., green roofs) and urban planning indicators

% TODO: different impacts in different cities \cite{zhou2017effects,gret2020urban}

\subsection*{Implications to urban planning}

The spatiotemporal patterns of LULC change observed during the last 40 years in the Lausanne agglomeration have been characterized by infilling development and a progressive coalescensce of the artificial surfaces in its inner ring \cite{bosch2020spatiotemporal}.
% Overall decrease of urban green space in European cities due to infill development \cite{haaland2015challenges}.
Such an infilling trend urges for careful evaluation of the beneficial ecosystem services provided by urban green spaces, which should be balanced against the adverse consequences of urban sprawl \cite{haaland2015challenges,artmann2019urban}.

The approach proposed in this study maps locations in the current urban fabric where the tree canopy cover can be increased.
While part of this urban greening might occur in impervious surfaces (e.g., in sidewalks, next to roads and in other impervious surfaces), most of the candidate locations currently correspond to urban green space (i.e., the ``garden'' LULC class).
Therefore, the potential heat mitigation suggested by the results study is not attainable in a scenario of severe infill development.
Additionally, densification strategies should consider that newly created patches of urban green space might result in less provision of ecosystem services than existing ones \cite{sun2017effects,wang2019quantity}.
Finally, infilling might exacerbate the unevenness of the accessibility to green areas by depriving dwellers of the most dense parts in city core from their few remaining urban green spaces.  % \cite{wolch2014urban}
% TODO: evenness of the distribution \cite{wolch2014urban}
% TODO: Spatially explicit is crucial
Spatial heterogeneity of this kind, which are encountered in many socioeconomic and environmental aspects of contemporary cities, are often hard to represent with aggregate indicators and highlight the importance of spatially explicit models to urban planning and decision making.

% TODO: human exposure, example heat wave 2003 in Switzerland \cite{grize2005heat}
The explicit representation of space is also crucial when considering the impacts of urban green space on human thermal comfort.
Although the simulated scenarios suggest that the impact of the spatial pattern of tree canopy on the air temperature is practically linear, the implications on human thermal comfort exhibit important thresholds.
% For example, the number of dwellers exposed to nighttime temperatures over 24$\degree$C can be reduced from 217051 to 31905, which respectively represents a 78.40 and 11.52\% of the total population in the study area.
For example, by increasing the tree canopy cover of a 25\% of the candidate pixels, the number of dwellers exposed to nighttime temperatures over 25$\degree$C can be reduced from 132641 to 69162, which respectively represents a 47.91 and 24.98\% of the total population in the study area.
Furthermore, the results suggest that such a population can be reduced by an additional 20000 dwellers by selecting such pixels to prioritize a spatial scattering of the tree canopy cover.
In Switzerland, the excess mortality associated to the heat wave of 2003 occurred over-proportionally to urban and sub-urban residents of its largest urban agglomerations \cite{grize2005heat}.
% TODO: relationship between heat and mortality in Switzerland \cite{ragettli2017exploring}
Furthermore, the association between temperature and mortality in extreme heat events in the largest Swiss urban agglomerations are exponential \cite{ragettli2017exploring}, which indicates that reducing temperatures by even fractions of a degree can have a dramatic impact on death rates.

\section*{Conclusion}
% TODO: conclusions
% The scenarios simulated in this study represent a new form of exploring the heat mitigation potential provided by the urban tree canopy cover.
% The results map locations where the tree canopy cover in the urban agglomeration of Lausanne can be increased, and shows that such changes can result in urban nighttime temperatures that are up to 2$\degree$C lower.
% Additionally, the number of dwellers exposed to nighttime temperatures over 24$\degree$C can be reduced from 217051 to 31905, which respectively represents a 78.40 and 11.52\% of the total population in the study area.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=.98\textwidth]{figures/monitoring-stations}
  \caption{\label{fig:monitoring-stations} Locations of the monitoring stations used to get the $T_{ref}$ and $UHI_{max}$ parameters. The basemap tile is provided by StamenDesign, under CC BY 3.0, with data from OpenStreetMap, under ODbL.}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-regplot}
  \caption{\label{fig:scenarios-prop-regplot} Relationship between the proportion of candidate pixels transformed and the average simulated temperature $\overline{T}$ for each scenario sample. The translucent bands around the regression line represent the 95\% confidence intervals estimated using a bootstrap.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=.6\textwidth]{figures/scenarios-prop-hists}
  \caption{\label{fig:scenarios-prop-hists} Histogram of raster temperature values for a 25, 50, 75 and 100\% of the candidate pixels transformed. The temperature rasters for each histogram are computed by averaging the 10 simulations with the same proportion of candidate pixels transformed.}
\end{figure}
